---
title: "Regression Models Project"
output: 
  pdf_document : default
  html_document:
    keep_md: TRUE
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = TRUE)
```

## Executive Summary
### Task to do :
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

- Is an automatic or manual transmission better for MPG?

- Quantify the MPG difference between automatic and manual transmissions

In this analysis we are attempting to find out whether a manual or automatic transmission is better for miles per gallon (mpg). This was done using a statistical analysis to quantify how different mpg is for cars using manual and automatic transmissions.
We will use the mtcars dataset.
Manual transmission will yield better miles per gallon, when compared with Automatic. On average, a manual car will achieve 24 mpg, versus 17 mpg for automatics.

## Exploring the dataset
```{r, echo=TRUE}
#Loading Dataset mtcars.
library(ggplot2)
library(datasets)
data("mtcars")

#Summarizing data.
head(mtcars)
str(mtcars)
summary(mtcars$mpg)
```

The dataset has 32 observations of 11 variables.

#### #Making am variable a factor variable which contains transmission data.
```{r, echo=TRUE}
mtcars$am <- as.factor(as.numeric(mtcars$am))
```

### Boxplot to show the difference between automatic and mannual transmission.
```{r, echo=TRUE}
plot1 <- ggplot(data = mtcars, aes(x = am, y = mpg, fill = am)) + 
         geom_boxplot() + xlab("Types of Transmission(0 - Automatic, 1- Mannual)") +
         ylab("Miles/(US) Gallon") + ggtitle("Relation b/w Transmission and Mpg")

plot1
```

### Splitting data according to type of transmission
```{r, echo=TRUE}
MotorTrend <- split.data.frame(mtcars, mtcars$am)
```

### Range of the mpg with different transmission.
```{r, echo=TRUE}
range(MotorTrend$`0`$mpg)
range(MotorTrend$`1`$mpg)
```

### Mean of the mpg with different transmission.
```{r, echo=TRUE}
mean(MotorTrend$`0`$mpg)
mean(MotorTrend$`1`$mpg)
```

### Standard Deviation of the mpg with different transmission.
```{r, echo=TRUE}
sd(MotorTrend$`0`$mpg)
sd(MotorTrend$`1`$mpg)
```

### Variance of the mpg with different transmission.
```{r,echo=TRUE}
var(MotorTrend$`0`$mpg)
var(MotorTrend$`1`$mpg)
```

### Performing t.test on data.
```{r,echo=TRUE}
t.test(MotorTrend$`0`$mpg, MotorTrend$`1`$mpg)
```

The boxplot above clearly indicates that manual transmissions provide better mileage than automatic. To test this, a test is done that rejects the null hypothesis. The transmission type is in fact significantly correlated to mileage. Regression analyses will now be performed to quantify how much of a factor transmission type accounts for mileage.

## Regression

### Regression model in which mpg depends on transmission.
```{r, echo=TRUE}
fit <- lm(mpg ~ factor(am), mtcars)
summary(fit)
```

The p-value is very low, so we will not reject the hypothesis. However, the R-squared value for this test is only = .36, suggesting that only a third or so of variance in MPG can be attributed to transmission type alone.

### Multivariable Regression Model
```{r, echo=TRUE}
fit2 <- lm(mpg  ~ . , mtcars)
summary(fit2)
```

From the above Analysis of Variance, we can look for p-values of less than .5. We can see cyl, wt, hp, show strong correlations and significance for the model. Hence we choose those variables plus am for a linear model. This gives us the following model below:

```{r, echo=TRUE}
fit3 <- lm(mpg ~ am + cyl + hp + wt, data = mtcars)
summary(fit3)
```

This Multivariable Regression test now gives us an R-squared value of over .8, suggesting that 83% or more of variance can be explained by the multivariable model. P-values for cyl (number of cylinders) and weight are below 0.5, suggesting that these are confounding variables in the relation between car Transmission Type and Miles per Gallon. 
This second provided the following results: a 1.48MPG increase from manual transmissions over automatic ones with the additional variables (multivariable) model explaining 85% of the MPG performance.

## Appendix

```{r, echo=TRUE}
par(mar = c(4,4,2,2), mfrow = c(2,2))
plot(fit3)
```

```{r,echo=TRUE}
par(mar = c(1,1,1,1))
pairs(mpg ~., mtcars)
```
