---
title: " Analyzing the ToothGrowth data in the R datasets package"
output: 
  pdf_document : default
  html_document:
    keep_md: TRUE
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = TRUE)
```

### 1. Load the ToothGrowth data and perform some basic exploratory data analyses.
#### Loading Libraries and dataset
```{r,echo=TRUE}
library(ggplot2)
library(datasets)
data("ToothGrowth")
```
### 2. Provide a basic summary of the data.
```{r, echo=TRUE}
dim(ToothGrowth)
summary(ToothGrowth)
unique(ToothGrowth$dose)
```
#### Making Boxplot of the Toothdata.
```{r echo=TRUE, fig.height=3, fig.width=3}
qplot(x = supp, y = len, data = ToothGrowth, facets = ~ dose, 
      main = "Tooth Growth by Supplement Type and Dosage", xlab="Supplement",
      ylab = "Length of Tooth") + geom_boxplot(aes(fill = supp))
```
**According to the plot there is a statistically significant difference between teeth length and dose levels across both delivery methods, as the dose increases so does length.**
#### Subsetting data according to Doses.
```{r,echo=TRUE}
Dose0.5 <- subset.data.frame(x = ToothGrowth, dose == 0.5) 
Dose1.0 <- subset.data.frame(x = ToothGrowth, dose == 1.0) 
Dose2.0 <- subset.data.frame(x = ToothGrowth, dose == 2.0)
```
### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
#### T.test on the data with 0.5 Dose
```{r, echo=TRUE}
T.TestDose0.5 <- t.test(len ~ supp, data = Dose0.5)
T.TestDose0.5
```
**p-value** when dosage is 0.5 is **0.006359** < 0.05.
#### T.test on the data with 1.0 Dose
```{r,echo=TRUE}
T.TestDose1.0 <- t.test(len ~ supp, data = Dose1.0)
T.TestDose1.0
```
**p-value** when dosage is 1.0 is **0.001038** < 0.05.
#### T.test on the data with 2.0 Dose
```{r,echo=TRUE}
T.TestDose2.0 <- t.test(len ~ supp, data = Dose2.0)
T.TestDose2.0
```
**p-value** when dosage is 2.0 is **0.9639** > 0.05.
#### T.test on the ToothGrowth Data when length depends on supplement. 
```{r, echo=TRUE}
T.TestToothGrowth <- t.test(len ~ supp, data = ToothGrowth)
T.TestToothGrowth
```
**p-value** when length depends on supplement is **0.06063** > 0.05.
### 4. State your conclusions and the assumptions needed for your conclusions.
In the previous section of this report we drew some conclusions from our tests.
**It appears that there is a significant relationship between teeth length and dose levels across both delivery methods, as the dose increases so does length. On the other hand, there doesn't seem to be a statistically significant difference between delivery methods, with OJ apparently more effective at dose levels 0.5 and 1, and VC slightly more effective at dose level 2**