---
title: "Developing Data Projects"
author: "Shubham Kumar"
date: "7/30/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Objective

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either github pages or Rpubs.
Your webpage must contain the date that you created the document and it must contain a plot created with Plotly.

## Slide with R code.

```{r, echo=TRUE}
library(ggplot2)
library(plotly)

data("iris")

str(iris)
```

## Plot with Relation between Species and its Sepal length.

```{r , echo = TRUE}
plot_ly(data = iris, y = ~Sepal.Length, color = ~Species, type = "scatter", mode = "lines")
```

## Plot with Relation between Species and its Petal length.

```{r , echo=TRUE}
plot_ly(data = iris, y = ~Petal.Length, color = ~Species, type = "scatter", mode = "lines")
```

## Plot with Relation between Species and its Sepal width.

```{r , echo=TRUE}
plot_ly(data = iris, y = ~Sepal.Width, color = ~Species, type = "scatter", mode = "lines")
```


## Plot with Relation between Species and its Petal width.

```{r , echo=TRUE}
plot_ly(data = iris, y = ~Petal.Width, color = ~Species, type = "scatter", mode = "lines")
```