---
title: "Practical-ML Project"
output: 
  html_document:
    keep_md: TRUE
  pdf_document : default
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = TRUE, cache = TRUE)
```

# Overview
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

The data consists of a Training data and a Test data (to be used to validate the selected model).

The goal of your project is to predict the manner in which they did the exercise. This is the “classe” variable in the training set. You may use any of the other variables to predict with.

Note: The dataset used in this project is a courtesy of “Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers’ Data Classification of Body Postures and Movements”.

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Loading libraries and setting seed.
```{r, echo=TRUE}
library(caret)
library(reshape2)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(12345)
```

# Getting, Cleaning and Exploring Data.
## Getting Data.
The next step is loading the dataset from the URL provided above. The training dataset is then partinioned in 2 to create a Training set (70% of the data) for the modeling process and a Test set (with the remaining 30%) for the validations. The testing dataset is not changed and will only be used for the quiz results generation.

```{r, echo=TRUE}
#Loading datasets.
TrainData <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
TestData  <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
dim(TrainData)
dim(TestData)
```

## Creating Partition with data.
```{r, echo=TRUE}
inTrain  <- createDataPartition(TrainData$classe, p = 0.70, list = FALSE)
TrainSet <- TrainData[inTrain, ]
TestSet  <- TrainData[-inTrain, ]
dim(TrainSet)
dim(TestSet)
```

Both created datasets have 160 variables. Those variables have plenty of NA, that can be removed with the cleaning procedures below. The Zero variance variables are also removed.

### Removing data with zero variance 

Since they are considered to have less predictive power. **(STACKEXCHANGE)**

```{r,echo=TRUE}
ZeroVariance <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -ZeroVariance]
TestSet  <- TestSet[, -ZeroVariance]
dim(TrainSet)
dim(TestSet)
```

## Removing Data which alters modelling.
```{r,echo=TRUE}
TrainSet <- TrainSet[,-(1:5)]
TestSet <- TestSet[,-(1:5)]
dim(TrainSet)
dim(TestSet)
```

**This was very neccessary step to do because if the columns not removed all the models will lead to accuracy = 1 even in the Tree model as the variables and data are very similar and at the end doing prediction of TestData all the prediction will be "A" and no other classe.**

### Removing columns with NA values.
```{r, echo=TRUE}
NAS <- sapply(TrainSet, function(x) mean(is.na(x))) > 0
TrainSet <- TrainSet[, NAS == FALSE]
TestSet  <- TestSet[, NAS == FALSE]
dim(TrainSet)
dim(TestSet)
```

### Checking data with high correlation.

A correlation among variables is analysed.

```{r, echo=TRUE}
Correlation <- cor(TrainData[sapply(TrainData, is.numeric)])
Correlation[is.na(Correlation)] <- 0
Cor <- findCorrelation(Correlation)
```

#### Correlated data.
```{r, echo=TRUE}
names(TrainSet)[Cor]
```

```{r,echo=TRUE}
Correlation <- melt(Correlation)
qplot(x = Var1, y = Var2, data = Correlation, fill = value, color = value)
```

The Light Blue Line represents that data is collinear.

**From the graph it shows there is a correlation in variables.**

# Modelling.

Three methods will be applied to model the regressions (in the Train dataset) and the best one (with higher accuracy when applied to the Test dataset).

A Confusion Matrix is plotted at the end of each analysis to better visualize the accuracy of the models.

# The methods are:

## Tree Model
```{r,echo=TRUE}
ModelTree <- rpart(classe ~ ., data = TrainSet, method="class")
fancyRpartPlot(ModelTree)
```

### Prediction with Tree Model
```{r,echo=TRUE}
PredictModelTree <- predict(ModelTree, newdata = TestSet, type = "class")
ConfMatTree <- confusionMatrix(PredictModelTree, reference = TestSet$classe)
ConfMatTree
```

**Accuracy of Tree Model is 73%**

## Random Forest
```{r,echo=TRUE}
ControlRF <- trainControl(method="cv", number = 3)
ModelRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=ControlRF)
ModelRandForest$finalModel
```

### Prediction with Random Forest Model.
```{r,echo=TRUE}
PredictRandForest <- predict(ModelRandForest, newdata = TestSet)
ConfMatRandForest <- confusionMatrix(PredictRandForest, TestSet$classe)
ConfMatRandForest
```

**Accuracy of Random Forest Model is 99.9%**

## GBR Model 
```{r,echo=TRUE}
GBR <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
ModelGBM  <- train(classe ~ ., data = TrainSet, method = "gbm", trControl = GBR
                   , verbose = FALSE)
ModelGBM$finalModel
```

### Prediction with GBR Model
```{r,echo=TRUE}
PredictModelGBM <- predict(ModelGBM, newdata = TestSet)
ConfMatGBM <- confusionMatrix(PredictModelGBM, TestSet$classe)
ConfMatGBM
```

**Accuracy of GBR(Generalized Boosted Regression) Model is 98.83**

**We will use Random Forest model to predict results of the TestData because it has very high accuracy than Tree and GBR Model.**

```{r, echo=TRUE}
PredictTest <- predict(ModelRandForest, newdata = TestData)
PredictTest
```