---
title: "Analysis of the Adverse Health and Economic Impacts of US Storms"
output: 
  html_document:
    keep_md : TRUE
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = TRUE)
```

## 1: Synopsis
The goal of the assignment is to explore the NOAA Storm Database and explore the effects of severe weather events on both population and economy.The database covers the time period between 1950 and November 2011.

The following analysis investigates which types of severe weather events are most harmful on:

1. Health (injuries and fatalities) 
2. Property and crops (economic consequences)

Information on the Data: [Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)

## 2: Data Processing

### 2.1: Data Loading

Loading Required packages and downloading the raw data file from the url and reading data directly to form data.table.
It requires R.utils package to extract bz2 files and to read files from it.
```{r DataLoading, echo=TRUE}
library(data.table)
library(ggplot2)
library(gridExtra)
library(grid)
library(R.utils)

path <- getwd()
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
stormdata <- data.table::fread(url)
```

### 2.2: Examining Column Names

```{r Columnnames and Head, echo=TRUE}
#Column names  and some values of the table.
names(stormdata)
head(stormdata)
```

### 2.3: Data Selection

Extracting only Neccessary Columns and Data.
```{r DataSubsetting, echo=TRUE}
stormdata <- stormdata[,c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG",   "PROPDMGEXP", "CROPDMG", "CROPDMGEXP")]
```

### 2.4: Converting Exponent Columns into Actual Exponents instead of (-,+, H, K, etc)

Making the PROPDMGEXP and CROPDMGEXP columns cleaner so they can be used to calculate Property and Crop cost and the values can be user readble.

```{r CorrectingExponents,echo=TRUE}
#Checking for distint values.
unique(stormdata$PROPDMGEXP)
unique(stormdata$CROPDMGEXP)

#Changing the values to user readable values. 
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "K"] <- 1000
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "M"] <- 1e+06
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == ""] <- 1
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "B"] <- 1e+09
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "m"] <- 1e+06
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "0"] <- 1
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "5"] <- 1e+05
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "6"] <- 1e+06
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "4"] <- 10000
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "2"] <- 100
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "3"] <- 1000
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "h"] <- 100
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "7"] <- 1e+07
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "H"] <- 100
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "1"] <- 10
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "8"] <- 1e+08

# Assigning '0' to invalid exponent stormdata
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "+"] <- 0
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "-"] <- 0
stormdata$PROPDMGEXP[stormdata$PROPDMGEXP == "?"] <- 0

#Changing class of PROPDMGEXP to numeric.
stormdata$PROPDMGEXP <- as.numeric(stormdata$PROPDMGEXP)

# Assigning values for the crop exponent stormdata 
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "M"] <- 1e+06
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "K"] <- 1000
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "m"] <- 1e+06
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "B"] <- 1e+09
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "0"] <- 1
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "k"] <- 1000
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "2"] <- 100
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == ""] <- 1

# Assigning '0' to invalid exponent stormdata
stormdata$CROPDMGEXP[stormdata$CROPDMGEXP == "?"] <- 0

#Changing class of CROPDMGEXP to numeric.
stormdata$CROPDMGEXP <- as.numeric(stormdata$CROPDMGEXP)

```

### 2.5: Calcuating Total Property and Crop  Damage Cost.

```{r TotalPropertyCropCost,echo=TRUE}
# Calculating the property damage value
stormdata$PROPDMGVAL <- stormdata$PROPDMG * stormdata$PROPDMGEXP

# calculating the crop damage value
stormdata$CROPDMGVAL <- stormdata$CROPDMG * stormdata$CROPDMGEXP
```

### 2.6:Calculating total numbers by events and selecting data with values greater than 0.

```{r Cleanata , echo=TRUE}
FatalitiesData <- aggregate(FATALITIES ~ EVTYPE, stormdata, FUN = sum, na.rm = TRUE)
FatalitiesData <- FatalitiesData[FatalitiesData$FATALITIES > 0,]

InjuriesData <- aggregate(INJURIES ~ EVTYPE, stormdata, FUN = sum, na.rm = TRUE)
InjuriesData <- InjuriesData[InjuriesData$INJURIES > 0,]

PropDmgValueData <- aggregate(PROPDMGVAL ~ EVTYPE, stormdata, FUN = sum, na.rm = TRUE)
PropDmgValueData <- PropDmgValueData[PropDmgValueData$PROPDMGVAL > 0,]

CropDmgValueData <- aggregate(CROPDMGVAL ~ EVTYPE, stormdata, FUN = sum, na.rm = TRUE)
CropDmgValueData <- CropDmgValueData[CropDmgValueData$CROPDMGVAL > 0,]

```

### 2.7: Ordering Data with Highest Values

```{r OrderedData, echo=TRUE}
#Ordering events with highest fatalities.
FatalitiesData <- FatalitiesData[order(-FatalitiesData$FATALITIES), ]

#Ordering events with highest injuries.
InjuriesData <- InjuriesData[order(-InjuriesData$INJURIES), ]

#Ordering events with highest Property Damage.
PropDmgValueData <- PropDmgValueData[order(-PropDmgValueData$PROPDMGVAL), ]

#Ordering events with highest injuries.
CropDmgValueData <- CropDmgValueData[order(-CropDmgValueData$CROPDMGVAL), ]
```

## 3: Results

### 3.1: Events with Highest Fatalities and Injuries.

```{r FatalitiesInjuries, echo=TRUE}
head(FatalitiesData)
head(InjuriesData)
```

### 3.2: Events with Highest Property Damage and Crop Damage.

```{r PropCropDamage, echo=TRUE}
head(PropDmgValueData)
head(CropDmgValueData)
```

### 3.3: Graph of Events with Highest Fatalitites and Injuries.

```{r Graph1, echo=TRUE}
plot1 <- ggplot(data = FatalitiesData[1:10,], aes(x = reorder(EVTYPE,                  FATALITIES), y = FATALITIES/1000, color = EVTYPE, fill = EVTYPE)) +
         geom_bar(stat = "identity")  + coord_flip() + 
         ylab("Total number of Fatalities in Thousands") + xlab("Event type")          + ggtitle("Fatalities in US by Top 10 Events")

#Plotting events with highest Injuries
plot2 <- ggplot(data = InjuriesData[1:10,], aes(x = reorder(EVTYPE, INJURIES),          y = INJURIES/1000, color = EVTYPE, fill = EVTYPE)) +
         geom_bar(stat = "identity") + coord_flip() +
         ylab("Total number of Injuries in Thousands") + xlab("Event type") +
         ggtitle("Injuries in US by Top 10 Events")

#Combining them to make 1 Plot
grid.arrange(plot1, plot2, nrow =2)
```

### 3.4: Graph of Events with Highest Property and Crop Damage.

```{r Plot2, echo=TRUE}
#Plotting events with highest Property Damage.
plot3 <- ggplot(data = PropDmgValueData[1:10,], aes(x = reorder(EVTYPE, PROPDMGVAL),y = PROPDMGVAL/1000000, color = EVTYPE, fill = EVTYPE)) + geom_bar(stat = "identity")  + coord_flip() + ylab("Total Damage to Property in  Million") + xlab("Event type") + ggtitle("Property Damage in US by Top 10 Events")

#Plotting events with highest Crop Damage.
plot4 <- ggplot(data = CropDmgValueData[1:10,], aes(x = reorder(EVTYPE, CROPDMGVAL), y = CROPDMGVAL/1000000, color = EVTYPE, fill = EVTYPE)) + geom_bar(stat = "identity")  + coord_flip() + ylab("Total Damage to Crop in Million") + xlab("Event type") + ggtitle("Crop Damage in US by Top 10 Events")

#Combining them to make 1 plot.
grid.arrange(plot3, plot4, nrow =2)
```